Precision line-by-line forward models for \'echelle spectra\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-08-12 paper/paper1/ms.tex
Edits from @astrocaroline printed paper draft on Aug 12, 2022

\begin{document}
\shorttitle{Transfer learning with blas\'e}
\shortauthors{Gully \& Morley}
\title{Precision line-by-line forward models for \'echelle spectra}

@@

The T6 ultracool dwarf 2MASS J08173001$-$6155158 was recently observed with IGRINS, revealing a rich spectroscopic atlas of molecules in its $T_\mathrm{eff} = 1060 \pm 50$ cloud-free atmosphere \citep{2022MNRAS.tmp.1421T}.  The sea of moleculaar lines are so rich as to blur entirely the notion of isolated lines.  Virtually all lines should instead be considered pseudo lines, computed using the opacity agglomeration technique known as ``correlated-$k$''.  The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ rotational broadening \citep{2022MNRAS.tmp.1421T}.  This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.  The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.

@@

\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$ blas\'e line represents its (neighbor-weighted) Equivalent Width (EW).  That means that blas\'e can automatically and quickly distill a massive high-banwidth spectrum into a catalog of line positions and EWs.  Even more, the catalog would have some resilience to rotaional broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.  Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.  Blas\'e does not have these limitations.  The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.  This approach may suit stellar abundance applications, such as Galactic archaeology and near-field cosmology.

@@

The scaffolding of \texttt{blas\'e} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.  In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges. Here we enumerate some planned or conceivable extensions.  We break up the themes into two categories: scientific extensions (this Section) and technical innovations (Section \ref{technicalInnovations}).  We order the presentation of these extensions by perceived scientific impact:

>>>>

\begin{document}
\shorttitle{Transfer learning with blas\'e}
\shortauthors{Gully \& Morley}
\title{Precision line-by-line forward models for \'echelle spectra\footnote{Open source code at \url{https://github.com/gully/blase}}}

@@

The T6 ultracool dwarf 2MASS J08173001$-$6155158 was recently observed with IGRINS, revealing a rich spectroscopic atlas of molecules in its $T_\mathrm{eff} = 1060 \pm 50$ cloud-free atmosphere \citep{2022MNRAS.tmp.1421T}.  The sea of molecular lines are so rich as to blur entirely the notion of isolated lines.  Many lines should instead be considered pseudo lines.  The boundaries of lines for 2MASS J08173001$-$6155158 become even more amorphous in the presence of its moderately high $v\sin{i} = 22.5\pm 0.5 ~ \text{km/s}$ rotational broadening \citep{2022MNRAS.tmp.1421T}.  This T6 IGRINS spectrum therefore represents an extreme end of a test-case, in which the underlying notions of \emph{blas\'e} go outside the comparatively safe assumptions in stellar spectra.  The application of blas\'e to 2MASS J08173001$-$6155158 can be viewed as conducting a deconvolution step simultaneously paired with line-by-line inference.

@@

\emph{``Equivalent Width machine''}-- The area under the curve of each $j^{th}$ blas\'e line represents its (neighbor-weighted) Equivalent Width (EW).  That means that blas\'e can automatically and quickly distill a massive high-bandwidth spectrum into a catalog of line positions and EWs.  Even more, the catalog would have some resilience to rotational broadening, line blending, and other friction points that hamper such an industrial-scale approach with existing tooling.  Current applications of Equivalent Widths rely on either isolated lines or---for multi-epoch applications---the assumption of non-variable neighboring lines.  Blas\'e does not have these limitations.  The interpretations of these EWs may have to be treated with care if intercompared with EWs obtained in the conventional way.  This approach may suit stellar abundance applications, such as Galactic archaeology and near-field cosmology.

@@

The scaffolding of \texttt{blas\'e} is designed in a way to promote extensibility, so many scientific questions can be written down in the flexible language presented in this paper.  In some cases the flexibility, precision, speed, and ease-of-use may unlock new approaches to long-standing astrophysical questions and practical challenges. Here we enumerate some planned or conceivable extensions.  We break up the themes into two categories: scientific extensions (this Section) and technical innovations (Section \ref{technicalInnovations}).
